\input{immersed-boundary}
\input{fluid-discretization}
\input{cell-forces}

\section{Geometry of reconstructed surfaces}\label{sec:surfaces}

From the previous section, we have analytic expressions for the Lagrangian force density,
$\F$. The RBC and platelet force models require first and second derivatives along their
surfaces. Additionally, The IB force spreading operation requires quadrature weights for
the surfaces. This section finishes where we left off by discussing the construction of
the necessary discrete linear operators and quadrature weights $\weight[A]$ through the
use of RBF-based methods.

\input{rbf-interpolation}
\input{rbf-quadrature}

\subsection{A note on the endothelium}
While our choice of force density for the endothelium does not require surface
reconstruction or quadrature weights, the methods described above are equally applicable.
The periodicity of $\domain$ makes the endothelium a topological torus. The 4-dimensional
torus, $\mathbb{T}$, with parametrization
\begin{equation*}
    \Xp(\qs) = \cos\q[1]\e_1 + \sin\q[1]\e_2 + \cos\q[2]\e_3 + \cos\q[2]\e_4,
\end{equation*}
for $\qs\in\sites=[0,\,2\pi)^2$, is orientation agnostic and homogeneous. It has surface
area $4\pi^2$ and Jacobian $\tilde{J} = 1$. It is, however, trivial to choose a set of
points on $\mathbb{T}$ that generate equal quadrature weights without the need to solve a
system like~\eqref{eq:quadrature-weights}--\eqref{eq:quadrature-constraint}. Any regular
grid suffices, but we generate a set of $N$ points, whether used for interpolation or
not, with the spiral
\begin{align*}
    \q^2_i &= 2\pi (i-1)/N, \\
    \q^1_i &= \modulo\left(\left\lceil\!\sqrt{N}\mskip\thinmuskip\right\rceil\theta^2_i,\,2\pi\right).
\end{align*}
Using distances between points
on $\mathbb{T}$ to construct RBFs allow for the interpolation of, and quadrature on, any
toroidal surface. However, care needs to be taken in applying the constructed discrete
operators to a ``flat'' torus, such as the endothelium. If, for example, $L$ is the
discrete analogue of $\L$ evaluated at sample site $\sample\qs$ and the interpolated
function $\psi(\qs)$ decomposes into
\begin{equation*}
    \psi(\qs) = \bar{\psi}(\qs) + \text{periodic part},
\end{equation*}
where $\bar{\psi}$ is known, aperiodic, and independent of time, then
\begin{equation*}
    \L\psi(\sample\qs)
    \approx \L\bar{\psi}(\sample\qs) + L\left(\data{\vec{\psi}} - \data{\vec{\bar{\psi}}}\right)
    := L\data{\vec{\psi}} + \epsilon,
\end{equation*}
where $\data{\vec{\psi}}$ and $\data{\vec{\bar{\psi}}}$ are vectors formed by evaluating
$\psi$ and $\bar{\psi}$, respectively, at each point in $\data\sites$. Computing
$\epsilon$ for each sample site can be done simultaneously, resulting in the vector
$\vec{\epsilon}$. A correction like $\vec{\epsilon}$ may be needed for each linear
operator.
